---
title: 'Life Expentancy Country Comparison'
author: "Doris Liu"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    toc: yes
    toc_float: yes
---

```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(gapminder)  # gapminder dataset
library(here)
library(janitor)
```


I explored the `gapminder` dataset that has data on life expectancy, population, and GDP per capita for 142 countries from 1952 to 2007. 

```{r}
glimpse(gapminder)

head(gapminder, 20) # look at the first 20 rows of the dataframe

```

To produce two graphs of how life expectancy has changed over the years for the *country* and the *continent* I come from, I created the `country_data` and `continent_data` with the code below.

```{r}
country_data <- gapminder %>% 
            filter(country == "China") # China is where I come from

continent_data <- gapminder %>% 
            filter(continent == "Asia")
```

First is a plot of life expectancy over time for the single country I chose, which is China. 

```{r, lifeExp_one_country}
plot1 <- ggplot(data = country_data, mapping = aes(x = year, y = lifeExp))+
 geom_point() +
 geom_smooth(se = FALSE) + NULL

plot1
```

A title added to the plot. 

```{r, lifeExp_one_country_with_label}
plot1<- plot1 +
 labs(title = "Plot of life expectancy over time for China",
     x = "Year",
     y = "Life Expectancy") + NULL

plot1
```

Second is a plot for all countries in the *continent* I come from, which is Asia.

```{r lifeExp_one_continent}
ggplot(continent_data, mapping = aes(x = year, y = lifeExp, colour= country, group =country))+
 geom_point() + 
 geom_smooth(se = FALSE) + NULL
```

The final analysis explores life expectancy over time for different continents.

```{r lifeExp_facet_by_continent}
ggplot(data = gapminder , mapping = aes(x = year, y = lifeExp, colour= continent))+
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_wrap(~continent) +
  theme(legend.position="none") + #remove all legends
  NULL
```

> Conclusions about life expectancy since 1952:  

In terms of trends, since 1952, there have been quicker increases in life expectancies of American and Asian countries, while this growth appears to be slower in Europe and Oceania. Life expectancies seem to have stayed unchanged since 1990 in Africa.  
When it comes to country distributions, life expectancies of different countries in Americas, Europe and Oceania are more similar to each other, while those of African and Aisan countries disperse.  

