---
title: 'MFA 2022 Pre-programme Assignment'
author: "Doris Liu"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    toc: yes
    toc_float: yes
---

```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(gapminder)  # gapminder dataset
library(here)
library(janitor)
```


## Task 1: Biography of Doris Liu

**Doris Liu**  
[LinkedIn](www.linkedin.com/in/xinyi-doris-liu)

**Doris Liu** *(Chinese name Xinyi Liu)* is from Shanghai, China. She is now pursuing a **Masters in Financial Analysis degree** at **London Business School** after graduating with a BS degree in Statistics at Fudan University in China. Being one with love for financial vehicles and a sense of digits, Doris has strong incentives to work on trading desks so she is now working hard to apply for **Sales and Trading/Markets** summer interships. 

**Three thing that makes Doris unique:**  
- Dancer: Jazz, Urban, Kpop  
- Harry Potter fan: Magic knowledge well-equipped  
- Strong analytical and coding skills: In Stats



## Task 2: `gapminder` country comparison

I explored the `gapminder` dataset that has data on life expectancy, population, and GDP per capita for 142 countries from 1952 to 2007. 

```{r}
glimpse(gapminder)

head(gapminder, 20) # look at the first 20 rows of the dataframe

```

To produce two graphs of how life expectancy has changed over the years for the *country* and the *continent* I come from, I created the `country_data` and `continent_data` with the code below.

```{r}
country_data <- gapminder %>% 
            filter(country == "China") # China is where I come from

continent_data <- gapminder %>% 
            filter(continent == "Asia")
```

First is a plot of life expectancy over time for the single country I chose, which is China. 

```{r, lifeExp_one_country}
plot1 <- ggplot(data = country_data, mapping = aes(x = year, y = lifeExp))+
 geom_point() +
 geom_smooth(se = FALSE) + NULL

plot1
```

A title added to the plot. 

```{r, lifeExp_one_country_with_label}
plot1<- plot1 +
 labs(title = "Plot of life expectancy over time for China",
     x = "Year",
     y = "Life Expectancy") + NULL

plot1
```

Second is a plot for all countries in the *continent* I come from, which is Asia.

```{r lifeExp_one_continent}
ggplot(continent_data, mapping = aes(x = year, y = lifeExp, colour= country, group =country))+
 geom_point() + 
 geom_smooth(se = FALSE) + NULL
```

The final analysis explores life expectancy over time for different continents.

```{r lifeExp_facet_by_continent}
ggplot(data = gapminder , mapping = aes(x = year, y = lifeExp, colour= continent))+
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_wrap(~continent) +
  theme(legend.position="none") + #remove all legends
  NULL
```

> Conclusions about life expectancy since 1952:  

In terms of trends, since 1952, there have been quicker increases in life expectancies of American and Asian countries, while this growth appears to be slower in Europe and Oceania. Life expectancies seem to have stayed unchanged since 1990 in Africa.  
When it comes to country distributions, life expectancies of different countries in Americas, Europe and Oceania are more similar to each other, while those of African and Aisan countries disperse.  


## Task 3: Brexit vote analysis

Have a quick glimpse at the 2016 Brexit vote data in the UK from [Elliott Morris](https://www.thecrosstab.com/).  
Variables:  
- The main outcome variable (or y) is `leave_share`, which is the percent of votes cast in favour of Brexit, or leaving the EU.   
- Each row is a UK [parliament constituency](https://en.wikipedia.org/wiki/United_Kingdom_Parliament_constituencies).

```{r load_brexit_data, warning=FALSE, message=FALSE}
brexit_results <- read_csv(here::here("data_project","brexit_results.csv"))

glimpse(brexit_results)
```

I plotted a histogram, a density plot, and the empirical cumulative distribution function of the leave % in all constituencies.

```{r brexit_histogram, warning=FALSE, message=FALSE}

# histogram
ggplot(brexit_results, aes(x = leave_share)) +
  geom_histogram(binwidth = 2.5) + 
  labs(title = "Histogram",
       subtitle="leave share in all constituencies",
       x = "leave share (leave %)",
       y = "count") + NULL


# density plot-- think smoothed histogram
ggplot(brexit_results, aes(x = leave_share)) +
  geom_density() +
  labs(title = "Density plot",
       subtitle="leave share in all constituencies",
       x = "leave share (leave %)",
       y = "density") 


# The empirical cumulative distribution function (ECDF) 
ggplot(brexit_results, aes(x = leave_share)) +
  stat_ecdf(geom = "step", pad = FALSE) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Empirical cumulative distribution function (ECDF)",
       subtitle="leave share in all constituencies",
       x = "leave share (leave %)",
       y = NULL) 
  


```

One common explanation for the Brexit outcome was fear of immigration and opposition to the EU's more open border policy. To check the relationship (or correlation) between the proportion of native born residents (`born_in_uk`) in a constituency and its `leave_share`, I got the correlation between the two variables. The correlation is almost 0.5, which shows that the two variables are positively correlated.

```{r brexit_immigration_correlation}
brexit_results %>% 
  select(leave_share, born_in_uk) %>% 
  cor()
```

The scatterplot between these two variables tells similar story. 

```{r brexit_immigration_plot}
ggplot(brexit_results, aes(x = born_in_uk, y = leave_share)) +
  geom_point(alpha=0.3) +
  
  # add a smoothing line, and use method="lm" to get the best straight-line
  geom_smooth(method = "lm") + 
  
  # use a white background and frame the plot with a black box
  theme_bw() +
  
  labs(title = "Scatterplot",
       subtitle="Correlation between the proportion of native born residents and leave share",
       x = "% of native born residents",
       y = "leave share (leave %)") +  NULL
```


> Conclusions about Brexit vote:  

The constituencies with higher proportions of native born residents show higher leave share, which means more native constituencies might fear more of immigration and thus they tend to leave EU because of its more open boarder policy.


## Task 4: Animal rescue incidents attended by the London Fire Brigade

The next dataset explored is [The London Fire Brigade](https://data.london.gov.uk/dataset/animal-rescue-incidents-attended-by-lfb) which attends a range of non-fire incidents ('special services', assistance to animals). 

```{r load_animal_rescue_data, warning=FALSE, message=FALSE}

url <- "https://data.london.gov.uk/download/animal-rescue-incidents-attended-by-lfb/8a7d91c2-9aec-4bde-937a-3998f4717cd8/Animal%20Rescue%20incidents%20attended%20by%20LFB%20from%20Jan%202009.csv"

animal_rescue <- read_csv(url,
                          locale = locale(encoding = "CP1252")) %>% 
  janitor::clean_names()


glimpse(animal_rescue)
```
First is quick counts of the number of incidents by year.

```{r, instances_by_calendar_year}

animal_rescue %>% 
  dplyr::group_by(cal_year) %>% 
  summarise(count=n())

animal_rescue %>% 
  count(cal_year, name="count")

```

Next is the analysis on number of incidents by animal group. 

```{r, animal_group_percentages}
animal_rescue %>% 
  group_by(animal_group_parent) %>% 
  
  #group_by and summarise will produce a new column with the count in each animal group
  summarise(count = n()) %>% 
  
  # mutate adds a new column; here we calculate the percentage
  mutate(percent = round(100*count/sum(count),2)) %>% 
  
  # arrange() sorts the data by percent. Since the default sorting is min to max and we would like to see it sorted
  # in descending order (max to min), we use arrange(desc()) 
  arrange(desc(percent))


animal_rescue %>% 
  
  #count does the same thing as group_by and summarise
  # name = "count" will call the column with the counts "count" ( exciting, I know)
  # and 'sort=TRUE' will sort them from max to min
  count(animal_group_parent, name="count", sort=TRUE) %>% 
  mutate(percent = round(100*count/sum(count),2))


```

> Something strange in this table:  

There are two categories, Cat and cat, which might be typo causing data ambiguity.


The final analysis is to look at the notional cost for rescuing each of these animals. Fix `incident_notional_cost` to change it into a number and calculate summary statistics for each animal group. 

```{r, parse_incident_cost,message=FALSE, warning=FALSE}

# what type is variable incident_notional_cost from dataframe `animal_rescue`
typeof(animal_rescue$incident_notional_cost)

# readr::parse_number() will convert any numerical values stored as characters into numbers
animal_rescue <- animal_rescue %>% 

  # we use mutate() to use the parse_number() function and overwrite the same variable
  mutate(incident_notional_cost = parse_number(incident_notional_cost))

# incident_notional_cost from dataframe `animal_rescue` is now 'double' or numeric
typeof(animal_rescue$incident_notional_cost)

```


```{r, stats_on_incident_cost,message=FALSE, warning=FALSE}

animal_rescue %>% 
  
  # group by animal_group_parent
  group_by(animal_group_parent) %>% 
  
  # filter resulting data, so each group has at least 6 observations
  filter(n()>6) %>% 
  
  # summarise() will collapse all values into 3 values: the mean, median, and count  
  # we use na.rm=TRUE to make sure we remove any NAs, or cases where we do not have the incident cos
  summarise(mean_incident_cost = mean (incident_notional_cost, na.rm=TRUE),
            median_incident_cost = median (incident_notional_cost, na.rm=TRUE),
            sd_incident_cost = sd (incident_notional_cost, na.rm=TRUE),
            min_incident_cost = min (incident_notional_cost, na.rm=TRUE),
            max_incident_cost = max (incident_notional_cost, na.rm=TRUE),
            count = n()) %>% 
  
  # sort the resulting data in descending order. You choose whether to sort by count or mean cost.
  arrange(desc(mean_incident_cost))

```

> Conclusions: 

My analysis in comparing the mean and the median for each animal group tells me that for most animal groups, mean incident costs are lower than median, which means for more than half of the incidents(positively skewed), it took smaller amount of costs compared with mean.  
The outliers are Squirrel, Ferret and Rabbit.


Exploration in the distribution of incident cost for each animal group is as follows: 

```{r, plots_on_incident_cost_by_animal_group,message=FALSE, warning=FALSE}

# base_plot
base_plot <- animal_rescue %>% 
  group_by(animal_group_parent) %>% 
  filter(n()>6) %>% 
  ggplot(aes(x=incident_notional_cost))+
  facet_wrap(~animal_group_parent, scales = "free")+
  theme_bw()

base_plot + geom_histogram()
base_plot + geom_density()
base_plot + geom_boxplot()
base_plot + stat_ecdf(geom = "step", pad = FALSE) +
  scale_y_continuous(labels = scales::percent)



```

> Conclusions about incident notional costs in rescuing animals:  

Boxplot best communicates the variability of the `incident_notional_cost` values because it can show the 25%, 75% quantiles and median.  
1. Horses, deers, cows, foxes and other wild or heavy livestock animals are more expensive to rescue, probably due to their size and thus longer time to rescue. According to standard deviation of incident notional costs and boxplots, the costs to rescue these mentioned large-size animals also vary a lot.  
2. For cats and dogs, incident costs are lower and less diverse. Costs are even lower and less diverse for hamster，squirrel, ferret and rabbit.  
3. This correlation between animal size, incident cost and spread of values is possibly explained by that large animals are harder to control and might need more labour, time and materials to rescue. 



Additional Information:  
-   Who did you collaborate with: None  
-   Approximately how much time did you spend on this problem set: 2.5 hours  
-   What, if anything, gave you the most trouble: Though I have used R before, I haven't used Rmarkdown so I spent some time learning and debugging.


