---
title: "Yield Curve inversion"
author: "Study group B3: Doris Liu"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
```


The yield curve is appearing more important these days since the Fed is discussing its tapering policy.

Every so often, we hear warnings from commentators on the "inverted yield curve" and its predictive power with respect to recessions. An explainer what a [inverted yield curve is can be found here](https://www.reuters.com/article/us-usa-economy-yieldcurve-explainer/explainer-what-is-an-inverted-yield-curve-idUSKBN1O50GA). If you'd rather listen to something, here is a great podcast from [NPR on yield curve indicators](https://www.podbean.com/media/share/dir-4zgj9-6aefd11)

In addition, many articles and commentators think that, e.g., [*Yield curve inversion is viewed as a harbinger of recession*](https://www.bloomberg.com/news/articles/2019-08-14/u-k-yield-curve-inverts-for-first-time-since-financial-crisis). One can always doubt whether inversions are truly a harbinger of recessions, and [use the attached parable on yield curve inversions](https://twitter.com/5_min_macro/status/1161627360946511873).


## Get the data

In our case we will look at US data and use the [FRED database](https://fred.stlouisfed.org/) to download historical yield curve rates, and plot the yield curves since 1999 to see when the yield curves flatten. There's further information from an article that explains the [yield curve is and its inversion can be found here](https://fredblog.stlouisfed.org/2018/10/the-data-behind-the-fear-of-yield-curve-inversions/). 

First, we will load the yield curve data file that contains data on the yield curve since 1960-01-01
```{r download_historical_yield_curve, warning=FALSE}
yield_curve <- read_csv(here::here("data_project", "yield_curve.csv"))
glimpse(yield_curve)
```

Our dataframe `yield_curve` has five columns (variables):

- `date`: already a date object
- `series_id`: the FRED database ticker symbol
- `value`: the actual yield on that date
- `maturity`: a short hand for the maturity of the bond
- `duration`: the duration, written out in all its glory!


According to [Wikipedia's list of recession in the United States](https://en.wikipedia.org/wiki/List_of_recessions_in_the_United_States), since 1999 there have been two recession in the US: between Mar 2001–Nov 2001 and between Dec 2007–June 2009. 
The yield curve seems to flatten before these recessions. Usually the flattening yield curve is an economic indicator of recessions. 
Since 1999, short-term (3 months) yield has been higher than the long term (10 years) debt for three time: Early 2001(2001 Recession), late 2006 to 2007(housing bubble crisis, subprime lending crisis) and late 2019(covid pandemic). 


## spread (10year - 3months)

To create our final plot, we do the following steps:

 - Calculate the spread (10year - 3months)
 - Plot the spread between 10 years and 3 months as a blue/red ribbon, blue for positive and red for negative
 - Superimpose recessions as the grey areas in our plot
 - Plot the rugs on x-axis

We get the recession data first.
```{r setup_US-recessions, warning=FALSE}

# get US recession dates after 1946 from Wikipedia 
# https://en.wikipedia.org/wiki/List_of_recessions_in_the_United_States

recessions <- tibble(
  from = c("1948-11-01", "1953-07-01", "1957-08-01", "1960-04-01", "1969-12-01", "1973-11-01", "1980-01-01","1981-07-01", "1990-07-01", "2001-03-01", "2007-12-01","2020-02-01"),  
  to = c("1949-10-01", "1954-05-01", "1958-04-01", "1961-02-01", "1970-11-01", "1975-03-01", "1980-07-01", "1982-11-01", "1991-03-01", "2001-11-01", "2009-06-01", "2020-04-30") 
  )  %>% 
  mutate(From = ymd(from), 
         To=ymd(to),
         duration_days = To-From)

recessions
```

We manipulate the data to calculate the up-and-down spreads.
```{r create spread, up and down, warning=FALSE}
yield_curve_recession <- yield_curve %>%
  filter(maturity=="3m"|maturity=="10y") %>%
  group_by(date)%>%
  pivot_wider(id_col=1, names_from = duration, values_from=value) %>%
  clean_names() %>%
  mutate(spread = x10_year_treasury_rate - x3_month_treasury_bill,
         up = ifelse(spread>0, spread, 0), #up gives the diffrence between 10y and 3m when 10y>3m
         down = ifelse(spread<0, spread, 0)) #down gives the diffrence between 10y and 3m when 10y<3m

```

```{r plot the spread (10year - 3months), out.width="200%",fig.width=40, fig.height=25}
yield_curve_recession %>%
  ggplot()+
  #plot rectangles for recessions
  geom_rect(data=recessions %>% filter(From>="1960-01-01"),
            aes(xmin=From,xmax=To,ymin=-Inf,ymax=Inf),colour="grey",alpha=0.2)+ 
  scale_x_date(date_breaks = "2 years", date_labels = "%Y")+
  #plot rug for x asis
  geom_rug(aes(x=date,colour=ifelse(x10_year_treasury_rate>=x3_month_treasury_bill,">=0","<0")),sides="b")+ 
  scale_colour_manual(values=c("#EC7063","#2874A6"),name="Actual vs Expected ", guide=FALSE)+
  geom_line(aes(x=date,y=spread,fill="black"))+
  #plot the areas in blue using up when 10y>3m
  geom_ribbon(aes(x=date,ymin=0,
                  ymax=up),alpha=0.4,fill="#2874A6")+ 
  #plot the areas in red using down when 10y<3m
  geom_ribbon(aes(x=date,ymin=down,
                  ymax=0),alpha=0.4,fill="#EC7063")+ 
  theme_bw()  +
  theme(legend.position = "none",
        plot.title =element_text(size=16, face='bold',hjust = 0,margin = margin(10,0,10,0)),
        plot.subtitle =element_text(size=16, hjust = 0), 
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=12),
        axis.ticks.x = element_line(),
        axis.ticks.y=element_line(),
        axis.title.x = element_text(size=16,face='bold'),
        axis.title.y = element_text(size=16,face='bold'),
        ) +
  labs(title = "Yield Curve Inversion: 10-year minus 3-month U.S. Treasury rates", 
       subtitle = "Difference in % points, monthly averages,
Grey areas representing recessions", 
       caption= "Source: St. Louis Federal Reserve Economic Database (FRED)",
       x="", y="Yield Spread (10year - 3months)") 


```

